### 2024-11-05 manuscript.tex
Fix spelling

Global storm-resolving models (GSRMs) are the next avenue of climate modelling.
Among them is the 5-km Icosahedral Nonhydrostatic Weather and Climate Model
(ICON). Due to the high resolution, parametrisations of convection and clouds
are avoided. Standard-resolution models have substantial cloud biases over the
Southern Ocean (SO), affecting radiation and sea surface temperature.  We
evaluated SO clouds in ICON and the ERA5 and MERRA-2 reanalyses.  The SO is
dominated by low clouds, which cannot be observed accurately from space due to
overlapping clouds, attenuation, and ground clutter.  Instead, we analysed
about 2400 days of lidar observations from 31 voyages and a station using a
ground-based lidar simulator.  ICON and the reanalyses underestimate the total
cloud fraction by about 10 and 20\%, respectively. ICON and ERA5 overestimate
the cloud occurrence peak at about 500 m, potentially explained by their
lifting condensation levels being too high.  The reanalyses strongly
underestimate near-surface clouds or fog.  MERRA-2 tends to underestimate cloud
occurrence at all heights. Unstable conditions are the most problematic for the
models. In daily cloud cover, ICON and the reanalyses tend to be about 1 and 2
oktas clearer, respectively. Compared to radiosondes, potential temperature is
accurate in all, but ICON is too unstable over the low-latitude SO and too
humid in the boundary layer. MERRA-2 is too humid at all heights. SO cloud
biases are a substantial issue in the GSRM, but are an improvement over the
lower-resolution reanalyses. Explicitly resolved convection and cloud processes
were not enough to address the model cloud biases.

@@

Increasing climate model resolution is a common way of improving model accuracy
\citep{mauritsen2022}. This is of course limited by the available computational
power and a tradeoff with increasing parametrisation complexity. Current
computational availability and acceleration from general-purpose computing on
graphics processing units (GPUs) is progressing to enable km-scale (also
called k-scale) Earth system models (ESMs) and coupled atmosphere--ocean
general circulation models (AOGCMs) in research conditions today and
operationally in the forthcoming years. Therefore, it represents a natural
advance in climate modelling.  Global storm-resolving models (GSRMs) are
emerging as a new front in the development of high-resolution global climate
models, with horizontal grid resolutions of about 2--8 km
\citep{satoh2019,stevens2019}. This is enough to resolve mesoscale convective
storms, but smaller-scale convective plumes and smaller-scale cloud structure
remain unresolved. At a 5-km scale, non-hydrostatic processes also become
important, and for this reason such models are generally non-hydrostatic. The
terms global cloud-resolving models or global convection-permitting/-resolving
models are also sometimes used interchangeably with GSRMs but imply that clouds
or convection are resolved explicitly, which is not entirely true for GSRMs, as
this would require an even higher horizontal resolution \citep{satoh2019}.
Representative of these efforts is the DYnamics of the Atmospheric general
circulation Modeled On Non-hydrostatic Domains (DYAMOND) project
\citep{stevens2019,dyamond}, which is an intercomparison of nine global GSRMs
over two 40-day time periods in summer (1 August -- 10 September 2016) and winter
(20 January -- 1 March 2020). A new one-year GSRM intercomparison is currently
proposed by \cite{takasuka2024}, with the hope of also evaluating the seasonal
cycle and large-scale circulation.  An alternative to using a computationally
costly GSRM is to train an artificial neural network on GSRM output and use it
for subgrid-scale clouds, as done with the GSRM ICON by \cite{grundner2022} and
\cite{grundner2023}.

@@

The data sources contained ceilometer observations captured by the Vaisala CL51
operating at a wavelenght of 910 nm, the Vaisala CT25K operating at 905 nm, and
the Lufft CHM 15k operating at 1064 nm, described in detail below (Sections
\ref{sec:cl51} and \ref{sec:chm15k}). A ceilometer is a low-power near-infrared
vertically pointing lidar principally designed to  measure cloud base, but they
also measure the full vertical structure of clouds as long as the laser signal
is not attenuated by thick clouds, which can be used to infer additional
information such as a cloud mask and cloud occurrence by height.

@@

Here, we use a free-running (i.e., \emph{not} nudged or using prescribed SST)
coupled GSRM simulation made for the purpose of climate projection.  nextGEMS
has so far produced four cycles of model runs. We used a Cycle 3 run
\emph{ngc3028} produced in 2023 \citep{nextgems2023a,nextgems2023b} for a model
time period of 20 January 2020 to 22 July 2025, of which we analysed the full
years 2021--2024 (inclusive). While a Cycle 4 run was available, we could not
use it due to the unavailability of the necessary variables. The horizontal
resolution of ngc3028 is about 5 km.  The model output is available on 90
vertical levels and 3-hourly instantaneous temporal resolution.  Unlike current
general circulation models (GCMs), the strom-resolving version of ICON does not
use convective and cloud parameterization but relies on explicit simulation of
convection and clouds on the model grid. While this makes the code development
simpler without having to rely on uncertain parameterizations, it can miss
smaller-scale clouds below the grid resolution.  Turbulence and cloud
microphysics are still parameterized in this model.

@@

ERA5 \citep{era5} is a reanalysis produced by the ECMWF.  It is based on a
numerical weather prediction model IFS version CY41R2.  The horizontal
resolution is 0.25° in latitude and longitude, which is about 28 km in the
North--South direction and 14 km in the East--West direction at 60°S.
Internally, the model uses 137 vertical levels. Here, we use output at 1-hourly
instantaeous time intervals, except for radiation quantities, which are
accumulations (from these we calculate daily means).  Vertically resolved
quantities are made available on 37 pressure levels.

@@

Fig. \ref{fig:cyclone-stability}b shows a geographical distribution of the
relatively stable and unstable conditions as determined by the LTS (Section
\ref{sec:cyclone-stability}). Relatively unstable conditions are prevalent
in the middle SO (50--65°S), which might be explained by the relatively cold
near-surface air overlying the relatively warm sea surface. Relatively
stable conditions are prevalent elsewhere over the SO. The distribution is also
less zonally symmetric than the cyclonic activity.  In the high-latitude SO,
the presence of sea ice might have substantial stabilising effect
\citep{knight2024}. The ERA5 renalysis is also substantially more stable than
ICON across the whole region.

@@

Let us consider the main limitations of the presented results. The spatial
coverage of our dataset does not include most parts of the Indian Ocean and
Pacific Ocean sectors of the SO. Even though climatological features of the SO
are typically relatively uniform zonally, variations exist, such as those
related to the Antarctic Peninsula and the southern tip of South America. The
voyages were mostly undertaken in the Austral summer months and only rarely in
the winter months, due to the poor accesibility of this region during winter.
Therefore, our results are mostly representative of summer and to a lesser
extent, spring and autumn conditions.

@@

We analysed a total of about 2400 days of lidar and 2300 radiosonde
observations from 31 voyages/campaigns and a subantarctic station, covering the
Atlantic, Australian, and New Zealand sectors of the SO over the span of 10
years. This dataset, together with the use of a ground-based lidar simulator,
provided a fairly comprehensive basis for evaluating SO cloud and thermodynamic
profile biases in the GSRM ICON and the ERA5 and MERRA-2 reanalyses. Our
analysis provides a unique evaluation perspective different from satellite
observations -- one that is more suitable for evaluating bounary layer clouds,
which are predominant in this region. Furthermore, we subsetted our dataset by
low and high latitude bands, cyclonic activity, and stability in order to
identify how these conditions relate to the biases.

>>>>

Global storm-resolving models (GSRMs) are the next avenue of climate modelling.
Among them is the 5-km Icosahedral Nonhydrostatic Weather and Climate Model
(ICON). Due to the high resolution, parameterizations of convection and clouds
are avoided. Standard-resolution models have substantial cloud biases over the
Southern Ocean (SO), affecting radiation and sea surface temperature.  We
evaluated SO clouds in ICON and the ERA5 and MERRA-2 reanalyses.  The SO is
dominated by low clouds, which cannot be observed accurately from space due to
overlapping clouds, attenuation, and ground clutter.  Instead, we analysed
about 2400 days of lidar observations from 31 voyages and a station using a
ground-based lidar simulator.  ICON and the reanalyses underestimate the total
cloud fraction by about 10 and 20\%, respectively. ICON and ERA5 overestimate
the cloud occurrence peak at about 500 m, potentially explained by their
lifting condensation levels being too high.  The reanalyses strongly
underestimate near-surface clouds or fog.  MERRA-2 tends to underestimate cloud
occurrence at all heights. Unstable conditions are the most problematic for the
models. In daily cloud cover, ICON and the reanalyses tend to be about 1 and 2
oktas clearer, respectively. Compared to radiosondes, potential temperature is
accurate in all, but ICON is too unstable over the low-latitude SO and too
humid in the boundary layer. MERRA-2 is too humid at all heights. SO cloud
biases are a substantial issue in the GSRM, but are an improvement over the
lower-resolution reanalyses. Explicitly resolved convection and cloud processes
were not enough to address the model cloud biases.

@@

Increasing climate model resolution is a common way of improving model accuracy
\citep{mauritsen2022}. This is of course limited by the available computational
power and a trade-off with increasing parameterization complexity. Current
computational availability and acceleration from general-purpose computing on
graphics processing units (GPUs) is progressing to enable km-scale (also
called k-scale) Earth system models (ESMs) and coupled atmosphere--ocean
general circulation models (AOGCMs) in research conditions today and
operationally in the forthcoming years. Therefore, it represents a natural
advance in climate modelling.  Global storm-resolving models (GSRMs) are
emerging as a new front in the development of high-resolution global climate
models, with horizontal grid resolutions of about 2--8 km
\citep{satoh2019,stevens2019}. This is enough to resolve mesoscale convective
storms, but smaller-scale convective plumes and smaller-scale cloud structure
remain unresolved. At a 5-km scale, non-hydrostatic processes also become
important, and for this reason such models are generally non-hydrostatic. The
terms global cloud-resolving models or global convection-permitting/-resolving
models are also sometimes used interchangeably with GSRMs but imply that clouds
or convection are resolved explicitly, which is not entirely true for GSRMs, as
this would require an even higher horizontal resolution \citep{satoh2019}.
Representative of these efforts is the DYnamics of the Atmospheric general
circulation Modeled On Non-hydrostatic Domains (DYAMOND) project
\citep{stevens2019,dyamond}, which is an intercomparison of nine global GSRMs
over two 40-day time periods in summer (1 August -- 10 September 2016) and winter
(20 January -- 1 March 2020). A new one-year GSRM intercomparison is currently
proposed by \cite{takasuka2024}, with the hope of also evaluating the seasonal
cycle and large-scale circulation.  An alternative to using a computationally
costly GSRM is to train an artificial neural network on GSRM output and use it
for subgrid-scale clouds, as done with the GSRM ICON by \cite{grundner2022} and
\cite{grundner2023}.

@@

The data sources contained ceilometer observations captured by the Vaisala CL51
operating at a wavelength of 910 nm, the Vaisala CT25K operating at 905 nm, and
the Lufft CHM 15k operating at 1064 nm, described in detail below (Sections
\ref{sec:cl51} and \ref{sec:chm15k}). A ceilometer is a low-power near-infrared
vertically pointing lidar principally designed to  measure cloud base, but they
also measure the full vertical structure of clouds as long as the laser signal
is not attenuated by thick clouds, which can be used to infer additional
information such as a cloud mask and cloud occurrence by height.

@@

Here, we use a free-running (i.e., \emph{not} nudged or using prescribed SST)
coupled GSRM simulation made for the purpose of climate projection.  nextGEMS
has so far produced four cycles of model runs. We used a Cycle 3 run
\emph{ngc3028} produced in 2023 \citep{nextgems2023a,nextgems2023b} for a model
time period of 20 January 2020 to 22 July 2025, of which we analysed the full
years 2021--2024 (inclusive). While a Cycle 4 run was available, we could not
use it due to the unavailability of the necessary variables. The horizontal
resolution of ngc3028 is about 5 km.  The model output is available on 90
vertical levels and 3-hourly instantaneous temporal resolution.  Unlike current
general circulation models (GCMs), the storm-resolving version of ICON does not
use convective and cloud parameterization but relies on explicit simulation of
convection and clouds on the model grid. While this makes the code development
simpler without having to rely on uncertain parameterizations, it can miss
smaller-scale clouds below the grid resolution.  Turbulence and cloud
microphysics are still parameterized in this model.

@@

ERA5 \citep{era5} is a reanalysis produced by the ECMWF.  It is based on a
numerical weather prediction model IFS version CY41R2.  The horizontal
resolution is 0.25° in latitude and longitude, which is about 28 km in the
North--South direction and 14 km in the East--West direction at 60°S.
Internally, the model uses 137 vertical levels. Here, we use output at 1-hourly
instantaneous time intervals, except for radiation quantities, which are
accumulations (from these we calculate daily means).  Vertically resolved
quantities are made available on 37 pressure levels.

@@

Fig. \ref{fig:cyclone-stability}b shows a geographical distribution of the
relatively stable and unstable conditions as determined by the LTS (Section
\ref{sec:cyclone-stability}). Relatively unstable conditions are prevalent
in the middle SO (50--65°S), which might be explained by the relatively cold
near-surface air overlying the relatively warm sea surface. Relatively
stable conditions are prevalent elsewhere over the SO. The distribution is also
less zonally symmetric than the cyclonic activity.  In the high-latitude SO,
the presence of sea ice might have substantial stabilising effect
\citep{knight2024}. The ERA5 reanalysis is also substantially more stable than
ICON across the whole region.

@@

Let us consider the main limitations of the presented results. The spatial
coverage of our dataset does not include most parts of the Indian Ocean and
Pacific Ocean sectors of the SO. Even though climatological features of the SO
are typically relatively uniform zonally, variations exist, such as those
related to the Antarctic Peninsula and the southern tip of South America. The
voyages were mostly undertaken in the Austral summer months and only rarely in
the winter months, due to the poor accessibility of this region during winter.
Therefore, our results are mostly representative of summer and to a lesser
extent, spring and autumn conditions.

@@

We analysed a total of about 2400 days of lidar and 2300 radiosonde
observations from 31 voyages/campaigns and a subantarctic station, covering the
Atlantic, Australian, and New Zealand sectors of the SO over the span of 10
years. This dataset, together with the use of a ground-based lidar simulator,
provided a fairly comprehensive basis for evaluating SO cloud and thermodynamic
profile biases in the GSRM ICON and the ERA5 and MERRA-2 reanalyses. Our
analysis provides a unique evaluation perspective different from satellite
observations -- one that is more suitable for evaluating boundary layer clouds,
which are predominant in this region. Furthermore, we subsetted our dataset by
low and high latitude bands, cyclonic activity, and stability in order to
identify how these conditions relate to the biases.
